Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images

### 2021-12-14 tex/paper.wecs/corrections_Aluisio/mudancas_texto_DEC2021.tex
some corrections from Aluisio

LINE 24 "to images available at larger and larger from time to time" ==> "to large images available over long periods".

@@

LINE 7 "not affected" ==> "unaffected"

LINE 18 "is" XXXXX

LINE 20 "In more details," XXXXX

LINES 21-22 "not only ... problems" XXXXXX

LINES 33-34 "In the light ... concepts" ==> "Given the aforementioned discussion and motivation"

LINE -2 "A basic" ==> "basic"

@@

LINE 8 'needed" ==> "desirable".

@@

LINE 1 "In details," XXXXX

LINE 2 "for ultra-high" ==> "for high and ultra-high"

LINE 7 ", where " XXXXX

LINE 9 "is that" ==> "is such that"

LINE 9 "what" ==> "which"

LINE 10 "unfeasible, but only" ==> "unfeasible. Moreover, only"

LINE -3 "could" ==> "should"

>>>>

* LINE 24 "to images available at larger and larger from time to time" ==> "to large images available over long periods".

@@

* LINE 7 "not affected" ==> "unaffected"

* LINE 18 "is" XXXXX

* LINE 20 "In more details," XXXXX

* LINES 21-22 "not only ... problems" XXXXXX

* LINES 33-34 "In the light ... concepts" ==> "Given the aforementioned discussion and motivation"

* LINE -2 "A basic" ==> "basic"

@@

* LINE 8 'needed" ==> "desirable".

@@

* LINE 1 "In details," XXXXX

* LINE 2 "for ultra-high" ==> "for high and ultra-high"

* LINE 7 ", where " XXXXX

* LINE 9 "is that" ==> "is such that"

* LINE 9 "what" ==> "which"

* LINE 10 "unfeasible, but only" ==> "unfeasible. Moreover, only"

* LINE -3 "could" ==> "should"

### 2021-12-14 tex/paper.wecs/wecs_tgrs.tex
some corrections from Aluisio

\IEEEPARstart{C}{hange} detection is an important task performed in remote sensing image that allows researchers and engineers to identify and evaluate modifications on land surfaces captured by multi-temporal satellite images. Analyzing problems such as deforestation \cite{barreto2016deforestation}, rapid urbanization \cite{ban2012multitemporal} and glacier melting \cite{scher2021mapping} are of great importance to study the dynamics of regions sensitive to climate changes and human activity. Furthermore, the increase on availability of satellite images in the past years raises the challenge of applying computationally cheap methods to images available at larger and larger from time to time. A review for change detection in multi-temporal remote sensing is given by \cite{ban2016change}.

@@

A special attention has been given for multi-temporal change detection using Synthetic Aperture Radar (SAR) images. 
Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}, \cite{bovolo2005detail}. 
Known to be not affected by weather, cloud and sunlight conditions, SAR images rise as an essential data source in change detection applications \cite{bovolo2005detail}.
Conversely, due to its acquisition architecture, the speckle noise typically affects the images obtained by SAR sensors and demands additional pre-processing before its use.

Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task.
In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and is allows an efficient computational treatment.

In more details, wavelet-based methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} not only in remote sensing problems thanks to wavelet capabilities in capturing multi-scale/resolution information. 
Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. 
Analysis of SAR images have been investigated under different approaches using wavelet methods, such as \cite{atto2012multidate,bouhlel2015multivariate,celik2009multiscale,cui2012statistical}. 
%%%RGN: não foi definido/motivado o que é screening
However, there are variable selection methods that have been successfully applied in high-dimensional statistical models but that are still novel ideas in the wavelet and change detection literature. We show that one of these methods has an interesting potential to provide good results even with simple algorithms.

In the light of presented discussions and motivations, using Wavelet and data screening concepts, we propose the Wavelet Energies Correlation Screening (WECS), a novel unsupervised multi-temporal change detection method for SAR images.
%
The main idea behind WECS is built on ultra-high dimensional feature screening for the wavelet coefficients \cite{fan2020statistical}. Such method is usually employed in high-dimensional regression models to reduce the problem's dimension by subsetting the available covariates in such a way that true covariates are among the chosen ones with high probability \cite{fan2008sure}. We show that by applying the feature screening idea on multi-temporal images, we obtain a fast and accurate method to cover change regions with good detection rates.

This paper is organized as follows:
A basic concepts and definitions are presented in Section~\ref{section_theroy}.
Section \ref{section_method} introduces the problem and presents the proposed method. We show WECS performance on simulated multi-temporal image data in Section \ref{secExpSimulated}. In Section \ref{secExpActual} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from December 26th 2015 to December 3rd 2017.  Section \ref{section_discussion} concludes this study.% paper with a discussion.

@@

Wavelet methods have been widely applied to analyze images in signal processing literature, specially for tasks such as signal denoising and compression \cite{mallat1998wavelet}. The most common way of describing wavelet representations is as a multi-resolution decomposition, where a signal is represented on approximation and detail coefficients, which provide coarse and finer details of the signal, respectively. In practice, the discrete wavelet transform of matrices (image) consists in applying low and high-pass convolution filters to its rows and columns \cite{mallat1989theory}. In case of smoothing, applying such low-pass filter $J$ times to rows and columns of a matrix $\mathcal{I}$, yields a smooth image $\vX$. The number $J$ is also called resolution level, a tuning parameter for wavelet smoothing. %This wavelet smoothing is employed in the next section as a denoising step for image analysis.

Although the wavelet smoothing on images embraces an initial step to analyze the data, the main goal is to find the spatial changes over time. 
The smoothed images still contain many coefficients that need to be evaluated simultaneously, which characterizes a high-dimensional problem with multiple time series corresponding to each location in space. 
Consequently, retaining only the most essential locations driving overall modifications across time is needed. 
This issue is recurrent in regression analysis when selecting the most relevant variables.
The feature screening technique is a particularly efficient method to identify relevant variables, especially when the number of candidates is large.

%Using wavelet smoothing on images consists in an initial step to analyze the data, but the main goal in the analysis is to find changes in space and at instants of time. The smoothed images still contain a large number of coefficients that need to be evaluated simultaneously, what characterizes a high-dimensional problem with multiple time series corresponding to each location in space. We need to employ a method that retains the most important locations driving overall modifications across time. This type of problem is recurrent in statistics, and particularly in regression analysis it is mainly recognized as a variable selection problem. Methods like variable selection are largely employed to identify the most important covariates in a study, and one method particularly efficient in identifying relevant variables when the number of candidates is very large is the feature screening technique.

In details, the feature screening is a method originally designed for ultra-high dimensional regression models \cite{fan2008sure}. In order to explain how it works, consider the usual linear regression framework where $\vy$ is a $n\times 1$ vector of observations from a response variable and $\{\vW_1 \cdots \vW_p\}$ is a $n\times p$ matrix with explanatory variables on its columns, which are used in a linear model $\vy = \sum_{i=1}^{p}\beta_i\vW_i + \vepsilon$, where $\beta_1,\ldots,\beta_p$ are unknown parameters and $\vepsilon$ is a zero mean random noise. The problem setup is that $p$ is much larger than $n$, what makes standard regression methods unfeasible, but only a handful of the available covariates are relevant for the model, i.e., have a nonzero corresponding parameter $\beta_i$. The feature screening idea consists in computing the sample correlation $\mathrm{corr}(\vy,\vW_i)$ among response and explanatory variables, and then selecting those covariates whose correlation are among the highest values. Under suitable conditions, such method is known to select a set containing all true covariates with high probability. In the image change detection problem, our idea is that a similar approach could be used to detect change locations by taking an overall change measure as response variable and local (pixel) measures as potential covariates.

>>>>

\IEEEPARstart{C}{hange} detection is an important task performed in remote sensing image that allows researchers and engineers to identify and evaluate modifications on land surfaces captured by multi-temporal satellite images. Analyzing problems such as deforestation \cite{barreto2016deforestation}, rapid urbanization \cite{ban2012multitemporal} and glacier melting \cite{scher2021mapping} are of great importance to study the dynamics of regions sensitive to climate changes and human activity. Furthermore, the increase on availability of satellite images in the past years raises the challenge of applying computationally cheap methods to large images available over long periods. A review for change detection in multi-temporal remote sensing is given by \cite{ban2016change}.

@@

A special attention has been given for multi-temporal change detection using Synthetic Aperture Radar (SAR) images. 
Some of the references already mentioned are examples of change detection methods applied to SAR images \cite{barreto2016deforestation,ban2012multitemporal,scher2021mapping,quin2014mimosa}, \cite{jia2018novel}, \cite{hou2014unsupervised}, \cite{bovolo2005detail}. 
Known to be unaffected by weather, cloud and sunlight conditions, SAR images rise as an essential data source in change detection applications \cite{bovolo2005detail}.
Conversely, due to its acquisition architecture, the speckle noise typically affects the images obtained by SAR sensors and demands additional pre-processing before its use.

Face with the complexity imposed by the high-dimensionality of multi-temporal datasets in addition to the presence of speckle noise, the change detection process using SAR images becomes a challenge task.
In this context, the use of Wavelets rises as a convenient approach once such technique is robust when dealing with noise data and allows an efficient computational treatment.

Wavelet-based methods present many advantages for a plethora of statistical applications \cite{vidakovic1999statistical} thanks to wavelet capabilities in capturing multi-scale/resolution information. 
Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. 
Analysis of SAR images have been investigated under different approaches using wavelet methods, such as \cite{atto2012multidate,bouhlel2015multivariate,celik2009multiscale,cui2012statistical}. 
%%%RGN: não foi definido/motivado o que é screening
However, there are variable selection methods that have been successfully applied in high-dimensional statistical models but that are still novel ideas in the wavelet and change detection literature. We show that one of these methods has an interesting potential to provide good results even with simple algorithms.

Given the aforementioned discussion and motivation, we propose the Wavelet Energies Correlation Screening (WECS), a novel unsupervised multi-temporal change detection method for SAR images.
%
The main idea behind WECS is built on ultra-high dimensional feature screening for the wavelet coefficients \cite{fan2020statistical}. Such method is usually employed in high-dimensional regression models to reduce the problem's dimension by subsetting the available covariates in such a way that true covariates are among the chosen ones with high probability \cite{fan2008sure}. We show that by applying the feature screening idea on multi-temporal images, we obtain a fast and accurate method to cover change regions with good detection rates.

This paper is organized as follows: basic concepts and definitions are presented in Section~\ref{section_theroy}.
Section \ref{section_method} introduces the problem and presents the proposed method. We show WECS performance on simulated multi-temporal image data in Section \ref{secExpSimulated}. In Section \ref{secExpActual} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from December 26th 2015 to December 3rd 2017.  Section \ref{section_discussion} concludes this study.% paper with a discussion.

@@

Wavelet methods have been widely applied to analyze images in signal processing literature, specially for tasks such as signal denoising and compression \cite{mallat1998wavelet}. The most common way of describing wavelet representations is as a multi-resolution decomposition, where a signal is represented on approximation and detail coefficients, which provide coarse and finer details of the signal, respectively. In practice, the discrete wavelet transform of matrices (image) consists in applying low and high-pass convolution filters to its rows and columns \cite{mallat1989theory}. In case of smoothing, applying such low-pass filter $J$ times to rows and columns of a matrix $\mathcal{I}$, yields a smooth image $\vX$. The number $J$ is also called resolution level, a tuning parameter for wavelet smoothing.

Although the wavelet smoothing on images embraces an initial step to analyze the data, the main goal is to find the spatial changes over time. 
The smoothed images still contain many coefficients that need to be evaluated simultaneously, which characterizes a high-dimensional problem with multiple time series corresponding to each location in space. 
Consequently, retaining only the most essential locations driving overall modifications across time is desirable. 
This issue is recurrent in regression analysis when selecting the most relevant variables.
The feature screening technique is a particularly efficient method to identify relevant variables, especially when the number of candidates is large.

The feature screening is a method originally designed for high and ultra-high dimensional regression models \cite{fan2008sure}. In order to explain how it works, consider the usual linear regression framework where $\vy$ is a $n\times 1$ vector of observations from a response variable and $\{\vW_1 \cdots \vW_p\}$ is a $n\times p$ matrix with explanatory variables on its columns, which are used in a linear model $\vy = \sum_{i=1}^{p}\beta_i\vW_i + \vepsilon$, $\beta_1,\ldots,\beta_p$ are unknown parameters and $\vepsilon$ is a zero mean random noise. The problem setup is such that $p$ is much larger than $n$, which makes standard regression methods unfeasible. Moreover, only a handful of the available covariates are relevant for the model, i.e., have a nonzero corresponding parameter $\beta_i$. The feature screening idea consists in computing the sample correlation $\mathrm{corr}(\vy,\vW_i)$ among response and explanatory variables, and then selecting those covariates whose correlation are among the highest values. Under suitable conditions, such method is known to select a set containing all true covariates with high probability. In the image change detection problem, our idea is that a similar approach should be used to detect change locations by taking an overall change measure as response variable and local (pixel) measures as potential covariates.
